---
title: "Class 13: Structure based drug discovery"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and clean HIV-Pr receptor structure

Here we will download the 1HSG PDB structure from the online database,remove the bound drug from Merk and write out a protein only and ligand only new set of PDB files


We will use the bio3d package functions `get.pdb()`, `read.pdb()`, `atom.select()`, and finally `write.pdb()` 

```{r}
library(bio3d)

# Download pdb file
file.name <- get.pdb("1hsg")


hiv <- read.pdb(file.name)

hiv # Quick summary of contents of pdb structure object
```

Q1: What is the name of the two non protein resid values in this structure? What does resid correspond to and how would you get a listing of all reside values in this structure?

Answer: HOH (127), MK1 (1)
Corresponds to the water and merk residues present in the model. 


# Preparing intial protein and ligand input files
```{r}
prot <- atom.select(hiv, "protein",value = TRUE)
lig <- atom.select(hiv, "ligand", value = TRUE)

write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")

```

# Section 2.3: Inspecting my docking results
```{r}
library(bio3d)

res <- read.pdb("all.pdbqt", multi = TRUE)

write.pdb(res, "results.pdb")
```

Assessing results using RMSD
```{r}
# res <- read.pdb("all.pdbqt", multi = TRUE)

ori <- read.pdb("ligand.pdbqt")

rmsd(ori, res)
```


## Section 3: Exploring the conformational dynamics of protiens

# 3.1 Normal Mode Analysis (NMA)
```{r}
library(bio3d)
pdb <- read.pdb("1HEL")
modes <- nma(pdb)
plot(modes, sse = pdb)
```


Visualize NMA Results
```{r}
mktrj(modes, mode = 7, file = "nma_7.pdb")
```























